Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-07-31 Paper/intro.tex
..

Time evolution of a chemically reacting system is largely dependent upon rate
constants associated with individual reactions. The rate constants are
typically assumed to exhibit a certain correlation with temperature (e.g.
Arrhenius-type). Hence, accurate specification of the rate-controlling
parameters is critical to the fidelity of simulations. However, in practical
applications, the parameters are either specified using expert knowledge or
estimated based on a regression-fit to a set of sparse and noisy data.
(\alennote{Any references for this?})
Consequently, the underlying uncertainty associated with the parameters and
therefore the simulation is typically ignored.

Vohra et al.~calibrated the Arrhenius diffusion parameters to quantify the
uncertainty in mixing rates in Zr/Al reactive multilayers using a Bayesian
framework in~\cite{Vohra:2014}. The framework accounted for prior knowledge,
measurement noise, and the discrepancy between model and experiments.
Additionally, in~\cite{Vohra:2017}, a design strategy was implemented to
determine optimal conditions for experiments in order to enhance the
information gain in the uncertain diffusion parameters.  Bayesian calibration
was accelerated in these efforts using a polynomial chaos~\cite{Ghanem:2003,
Xiu:2002} surrogate for the model. However, the number of uncertain parameters
for calibration were limited to two in both cases. In higher dimensions, the
construction of surrogate as well as the calibration process become remarkably
challenging due to increase in computational effort.~\alennote{I don't see why
these references are being cited here. Are these cited as examples of efforts
in calibrating chemical systems?}

In this paper, we focus on mitigating the computational challenge
posed by uncertainty quantification (UQ) for applications involving a large
number of chemical reactions. We accomplish this by implementing
an efficient approach that aims to reduce the dimensionality of the problem. This is
done by
computing the so-called active subspace~\cite{Constantine:2015}.  The active
subspace, discussed later in~\ref{sub:ac}, constitutes the dominant
eigenspace of a matrix derived from the gradient information of the model
output with respect to an input. The dominant eigenspace essentially comprises
``important'' directions in the parameter space that predominantly capture the
variability in the model output due to uncertainty in the inputs.

>>>>

Time evolution of a chemically reacting system is largely dependent upon rate
constants associated with individual reactions. The rate constants are
typically assumed to exhibit a certain correlation with temperature (e.g.
Arrhenius-type). Hence, accurate specification of the rate-controlling
parameters is critical to the fidelity of simulations. However, in practical
applications, the parameters are either specified using expert knowledge or
estimated based on a regression-fit to a set of sparse and noisy
data~\cite{Burnham:1987, Burnham:1988, Vohra:2011, Sarathy:2012}.
Consequently, the underlying uncertainty associated with the parameters and
therefore the simulation is typically ignored. Nevertheless, the inclusion of
uncertainty due to corrupted calibration data, parametric uncertainty,
and the discrepancy between predictions and experiments
due to model inadequacy has gained prominence in recent
years~\cite{Vohra:2014, Vohra:2017, Morrison:2018, Hantouche:2018}. 
However, in complex mechanisms involving a large number of reactions,
characterizing the propagation of uncertainty from a large set of inputs to
the model output becomes remarkably challenging due to associated
computational effort. 
A major focus of this paper pertains to the implementation of a robust
framework that aims to identify `important' directions in the input space
that predominantly captures the variability in the model output. These
directions, referred to as the \textit{active subspaces}~\cite{Constantine:2015},
constitute the dominant
eigenspace of a matrix derived from the gradient information of the model
output with respect to an input. The active subspace methodology thus
focuses on reducing the dimensionality of the problem.
